---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Fanyi Zeng"
date: "02/06/22"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
load("C:/Users/zengf/Desktop/Data Science/Lab4/dennys.rda")
load("C:/Users/zengf/Desktop/Data Science/Lab4/laquinta.rda")
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

denny's dataset
dimensions: 6 columns and 1643 rows 
each row represents a single denny's store
each column represents a variable - (street) address, city, state, zip, longitude, latitude

```{r dn-dimensions}
ncol(dennys)
nrow(dennys)
head(dennys)
```

### Exercise 2

la quinta's dataset
dimensions: 6 columns and 909 rows 
each row represents a single la quinta hotel
each column represents a variable - (street) address, city, state, zip, longitude, latitude

```{r lq-dimensions}
ncol(laquinta)
nrow(laquinta)
head(laquinta)
```

### Exercise 3

All Denny's locations are within the country. Let's add a country variable for Denny's.

```{r dennys-states}
dennys %>%
  filter(!(state %in% states$abbreviation))
```

```{r dennys-country}
dennys <- dennys %>%
  mutate(country = "United States")
```


### Exercise 4

La Quinta have locations outside of US, including Mexico, Canada, China, New Zealand, Honduras, Turkey, United Arab Emirates, Chile, and Colombia.

```{r laquinta-states}
laquinta %>%
  filter(!(state %in% states$abbreviation))
```

### Exercise 5

Let's add a country variable for la quinta.

```{r laquinta-country}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico",
    state == "FM" ~ "Honduras"))
```

### Exercise 6

Let's filter out non-US locations.

```{r filter-nonUS}
laquinta <- laquinta %>%
  filter(country == "United States")
```

### Exercise 7

CA has the most Denny's locations.
TX has the most La Quinta's locations.

```{r most-state}
dennys %>%
  count(state) %>%
  arrange(desc(n))
laquinta %>%
  count(state) %>%
  arrange(desc(n))
```

Per square thousand miles:
DC has the most Denny's locations per 1000 sq miles.
RI has the most La Quinta's locations per 1000 sq miles.

```{r most-concentrated}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(concentration = n/area*1000) %>%
  arrange(desc(concentration))


laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(conc = n/area*1000) %>%
  arrange(desc(conc))
```

### Excersie 8

Let's plot the locations for both establishments in the U.S. and then in NC. We can see that not every La Quinta has a Denny's nearby. However, it seems to hold for TX.

```{r bind-and-plot}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
dn_lq <- bind_rows(dennys, laquinta)
ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point() +
  labs(title = "locations of both establishments", subtitle = "by latitude and longitude")
dn_lq %>%
  filter(state == "NC") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.5)) +
  geom_point() +
  labs(title = "locations of both establishments in NC", subtitle = "by latitude and longitude")
dn_lq %>%
  filter(state == "TX") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "locations of both establishments in TX", subtitle = "by latitude and longitude")
```
